# Acknowledging Previous Findings: Informed Priors

Setup:
```{r}
pacman::p_load(tidyverse, 
       glue,
       data.table,
       dplyr,
       moments,
       tidybayes,
       ggplot2,
       ggridges,
       plyr,
       ellipse,
       brms,
       cowplot,
       viridis,
       cmdstanr,
       ggrepel)
d <- read.csv('vowel_space_area_data.csv')
```

## Hands-on Exercises

### How to Encode Information in Priors
In this part, we'll have a closer look at how to encode information from previous studies within the prior specifications of our models. Let's start with the meta-analytic prior (i.e., Mean ES: 0.62, SE = 0.13) and add it to our prior specifications for the multi-level model. We do this by specifying the prior for the slope as normal(0.62, 0.13), as in the code block below:

```{r, results="hide", message=FALSE, warning=FALSE}
Articulation_f3 <- bf(ArticulationS ~ 1 + Register + (1+Register|Subject))
meta_analytic_prior <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, coef = Intercept, group = Subject),
  prior(normal(0.62, 0.13), class = b),
  prior(normal(0, 1), class = sd, coef = RegisterIDS, group = Subject),
  prior(normal(1, 0.5), class = sigma),
  prior(lkj(2), class = cor))
Articulation_MAprior_m3 <- 
  brm(
    Articulation_f3,
    data = d,
    save_pars = save_pars(all = TRUE),
    family = gaussian,
    prior = meta_analytic_prior,
    file = "Articulation_MAprior_m3",
    #refit = "on_change",
    sample_prior = T,
    iter = 10000, 
    warmup = 1000,
    cores = 2,
    chains = 2,
    #backend = "cmdstanr",
    #threads = threading(2),
    control = list(
      adapt_delta = 0.999,
      max_treedepth = 20))
```

```{r, warning=FALSE}
pp_check(Articulation_MAprior_m3, ndraws = 50)
plot(conditional_effects(Articulation_MAprior_m3), points = T)
summary(Articulation_MAprior_m3)
summary(Articulation_m3)
```

Q8: How does the slope estimate for this model with the meta-analytic prior compare to that with the skeptical prior (i.e., Articulation_m3)?

Answer: 
Articulation_m3 slope estimate: pop level = -0.39
Articulation_MAprior_m3 estimate: pop level = 0.43

The meta-analytically informed prior gives a model finding hyperarticulation in IDS, whereas the sceptical prior gives a model finding hypoarticulation. 
__________________________________________________________________________________________________________________________________________


Let's try to do the same analysis for the Danish prior (i.e., Mean ES: -0.1, SE = 0.04):

```{r, results="hide", message=FALSE}
danish_prior <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = sd, coef = Intercept, group = Subject),
  prior(normal(-0.1, 0.04), class = b),
  prior(normal(0, 1), class = sd, coef = RegisterIDS, group = Subject),
  prior(normal(1, 0.5), class = sigma),
  prior(lkj(2), class = cor))
Articulation_Danishprior_m3 <- 
  brm(
    Articulation_f3,
    data = d,
    save_pars = save_pars(all = TRUE),
    family = gaussian,
    prior = danish_prior,
    file = "Articulation_Danishprior_m3",
    #refit = "on_change",
    sample_prior = T,
    iter = 10000, 
    warmup = 1000,
    cores = 2,
    chains = 2,
    #backend = "cmdstanr",
    #threads = threading(2),
    control = list(
      adapt_delta = 0.999,
      max_treedepth = 20))
```

```{r}
pp_check(Articulation_Danishprior_m3, ndraws = 50)
plot(conditional_effects(Articulation_Danishprior_m3), points = T)
summary(Articulation_Danishprior_m3)
```

Q9: How does the slope estimate for this model with the Danish prior compare to that with the skeptical prior (i.e., Articulation_m3)?

Answer: 
Articulation_m3 slope estimate: pop level = -0.39
Articulation_Danishprior_m3 estimate: pop level = -0.11

With the prior informed by the Danish study, the model estimates hypoarticulation as with the sceptical prior, however to a much smaller extent than what the data with only the sceptical prior suggests. 
__________________________________________________________________________________________________________________________________________

### Prior-Posterior Updates Plot

Let's run the following code to extract and visualise the prior and posterior distributions from the different models:

```{r, warning=FALSE}
danish_prior_posterior <- as_draws_df(Articulation_Danishprior_m3) %>%
  mutate(priors = "Danish") %>%
  select(prior_b, b_RegisterIDS, priors)
  
ma_prior_posterior <- as_draws_df(Articulation_MAprior_m3) %>%
  mutate(priors = "MA estimates") %>%
  select(prior_b, b_RegisterIDS, priors)
  
skeptical_prior_posterior <- as_draws_df(Articulation_m3) %>%
  mutate(priors = "Skeptical estimates") %>%
  select(prior_b, b_RegisterIDS, priors)
Posterior <- rbind(danish_prior_posterior, ma_prior_posterior, skeptical_prior_posterior)
plot1 <- ggplot(Posterior) + 
  theme_classic() +
  ggtitle("Priors") +
  geom_density(aes(x = prior_b, fill = priors), alpha = 0.7) +
  xlim(c(-1.5, 1.5)) +
  geom_vline(xintercept = 0.0, linetype = 3) +
  scale_fill_manual(name = "Prior",
                  labels = c('Meta-analytic', "Danish", "Skeptical"),
                  values=c("#FC4E07", "steelblue", "#228B22")) +
  theme(plot.title = element_text(hjust = 0.5, size=15),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y=element_blank(),
        axis.line.x = element_line(size=0.1),
        axis.line.y = element_line(size=0.0),
        axis.text.x = element_blank(),
        axis.title.x=element_blank(),
        legend.position = "none")
plot2 <- ggplot(Posterior) + 
  theme_classic() +
  xlim(c(-1.5, 1.5)) +
  geom_density(aes(x = b_RegisterIDS, fill = priors), alpha = 0.7) +
  xlab('Effect Size') +
  ggtitle(expression(paste("Posteriors"))) +
  geom_vline(xintercept = 0.0, linetype = 3) +
  scale_fill_manual(name = "Priors:",
                  labels = c('Danish', 'Meta-analytic', "Skeptical"),
                  values=c("#FC4E07", "steelblue", "#228B22")) +
  scale_color_manual(values=c("#FC4E07","#228B22", "steelblue")) + 
  theme(plot.title = element_text(hjust = 0.5, size=15),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y=element_blank(),
        axis.line = element_line(size=0),
        legend.position = "bottom")
priors_posteriors_plot <- plot_grid(plot1, plot2, ncol=1)
priors_posteriors_plot
```


Q10: Why does the model with the Danish study prior not shift as much as those with the skeptical and meta-analytic priors?

Answer: It much more certain to begin, estimate has a low sd meaning that strong evidence from the data is needed to move the posterior estimate substantially. 


__________________________________________________________________________________________________________________________________________


Q11: What does this tell us about the nature of evidence accumulation in Bayesian models?

Answer: The more certain the distribution, be it prior or likelihood, the more influence it has on the posterior distribution. If we have a very certain prior, then data has little effect, and conversely, if we have data with a strong unambiguous trend, then changing the prior has little effect


__________________________________________________________________________________________________________________________________________
